--- 
title: '[Tree-Based Models](https://learn.datacamp.com/courses/tree-based-models-in-r)'
author: '[Kayla Friend](https://kaylafriend.github.io/)'
date: 'Last compiled: `r format(Sys.time(), "%b %d, %Y")`'
bibliography:
- book.bib
- packages.bib
description: Bookdown based on the datacamp course Tree-Based Models in R.
  The output format for this example is bookdown::gitbook.
documentclass: book
link-citations: yes
site: bookdown::bookdown_site
biblio-style: apalike
always_allow_html: true
---

# Prerequisites

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE, echo=FALSE}

```

This material is from the [DataCamp](https://www.datacamp.com) course [Tree-Based Models](https://learn.datacamp.com/courses/tree-based-models-in-r) by Erin L. and Gabriela de Queiroz.  Before using this material, the reader should have completed and be comfortable with the material in the DataCamp module [Tree-Based Models](https://learn.datacamp.com/courses/tree-based-models-in-r).


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
